---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# regionReport

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/regionReport.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/regionReport)
[![BioC dev status](http://www.bioconductor.org/shields/build/devel/bioc/regionReport.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/regionReport)
[![R build status](https://github.com/leekgroup/regionReport/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/leekgroup/regionReport/actions)
[![Support site activity, last 6 months: tagged questions/avg. answers per question/avg. comments per question/accepted answers, or 0 if no tagged posts.](http://www.bioconductor.org/shields/posts/regionReport.svg)](https://support.bioconductor.org/t/regionReport/)
[![GitHub issues](https://img.shields.io/github/issues/leekgroup/regionReport)](https://github.com/leekgroup/regionReport/issues)
<!-- badges: end -->

Generate HTML reports for a set of regions such as those from 
`r BiocStyle::Biocpkg('derfinder')` results or any other pipeline that defines a set of genomic regions.

Check the documentation for `derfinderReport()` for an example on how to create 
the necessary input files and generating the HTML report for `r BiocStyle::Biocpkg('derfinder')`
results. Or use:

```{r, eval = FALSE}
example("derfinderReport", "regionReport", ask = FALSE)
```

Similarly, check `renderReport()` for an example of a general report, or use:

````{r, eval = FALSE}
example("renderReport", "regionReport", ask = FALSE)
```

For `r BiocStyle::Biocpkg('DESeq2')` or `r BiocStyle::Biocpkg('edgeR')` results check `DESeq2Report()` and `edgeReport()`.

## Documentation

For more information about `derfinderPlot` check the vignettes [through Bioconductor](http://bioconductor.org/packages/regionReport) or at the [documentation website](http://leekgroup.github.io/regionReport).

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `regionReport` from [Bioconductor](http://bioconductor.org/) using the following code:

```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("regionReport")
```

## Citation

Below is the citation output from using `citation('regionReport')` in R. Please 
run this yourself to check for any updates on how to cite __regionReport__.

```{r 'citation', eval = requireNamespace('regionReport')}
print(citation("regionReport"), bibtex = TRUE)
```

Please note that the `regionReport` was only made possible thanks to many other R and bioinformatics software authors, which are cited either in the vignettes and/or the paper(s) describing this package.

## Code of Conduct
  
Please note that the regionReport project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Development tools

* Continuous code testing is possible thanks to [GitHub actions](https://www.tidyverse.org/blog/2020/04/usethis-1-6-0/)  through `r BiocStyle::CRANpkg('usethis')`, `r BiocStyle::CRANpkg('remotes')`, `r BiocStyle::Githubpkg('r-hub/sysreqs')` and `r BiocStyle::CRANpkg('rcmdcheck')` customized to use [Bioconductor's docker containers](https://www.bioconductor.org/help/docker/) and `r BiocStyle::Biocpkg('BiocCheck')`.
* Code coverage assessment is possible thanks to [codecov](https://codecov.io/gh) and `r BiocStyle::CRANpkg('covr')`.
* The [documentation website](http://leekgroup.github.io/derfinderPlot) is automatically updated thanks to `r BiocStyle::CRANpkg('pkgdown')`.
* The code is styled automatically thanks to `r BiocStyle::CRANpkg('styler')`.
* The documentation is formatted thanks to `r BiocStyle::CRANpkg('devtools')` and `r BiocStyle::CRANpkg('roxygen2')`.

For more details, check the `dev` directory.

This package was developed using `r BiocStyle::Biocpkg('biocthis')`.
