---
output:
  html_document:
    toc: true
    theme: united
  knitrBootstrap::bootstrap_document:
    theme.chooser: TRUE
    highlight.chooser: TRUE
---


<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Introduction to regionReport}
-->

HTML reports for a set of regions
=================================

If you wish, you can view this vignette online [here](http://lcolladotor.github.io/regionReport/).


```{r vignetteSetup, echo = FALSE, message = FALSE, warning = FALSE}
## Track time spent on making the vignette
startTimeVignette <- Sys.time()

## Bib setup
library('knitcitations')

## Load knitcitations with a clean bibliography
cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html')
# Note links won't show for now due to the following issue
# https://github.com/cboettig/knitcitations/issues/63

## Write bibliography information
write.bibtex(c(knitcitations = citation('knitcitations'), 
    derfinder = citation('derfinder'), 
    regionReport = citation('regionReport'),
    knitrBootstrap = citation('knitrBootstrap'),
    ggbio = citation('ggbio'),
    ggplot2 = citation('ggplot2'),
    rCharts = citation('rCharts'), 
    knitr = citation('knitr')[3],
    rmarkdown = citation('rmarkdown')),
    file = 'derfinderPlotRef.bib')
bib <- read.bibtex('derfinderPlotRef.bib')

## Fix some names to work with CRAN and GitHub versions
names(bib)[names(bib) == 'hester2013knitrbootstrap'] <- 'hester2014knitrbootstrap'
```


`regionReport` `r citep(bib[['colladotorres2014regionreport']])` creates HTML reports 
styled with `knitrBootstrap` `r citep(bib[['hester2014knitrbootstrap']])` for 
a set of regions such as `derfinder` `r citep(bib[['colladotorres2014derfinder']])` results.


Currently, this package includes a basic exploration analysis of `derfinder` 
results which we expect users to be interested in reproducing with their own
data. The analysis is written in [R Markdown](http://www.rstudio.com/ide/docs/r_markdown)
format and `derfinderReport()` takes the results from 
`derfinder`, performs a couple of setup operations, and then relies on
`knitr` `r citep(bib[['xie2014knitr']])`, `rmarkdown` 
`r citep(bib[['allaire2014rmarkdown']])`, and `knitrBootstrap` 
`r citep(bib[['hester2014knitrbootstrap']])` for generating the report. 

Other goodies in the report are powered by `ggbio` `r citep(bib[['yin2012ggbio']])`, 
`ggplot2` `r citep(bib[['wickham2009ggplot2']])`, and `rCharts`
`r citep(bib[['vaidyanathan2013rcharts']])`.

# Using `regionReport`

## General case

This is currently under construction.

## `derfinder` case

### Run `derfinder`

Prior to using `regionReport::derfinderReport()` you must use `derfinder` to analyze a specific
data set. While there are many ways to do so, we recommend using 
__analyzeChr()__ with the same _prefix_ argument. Then merging the results with 
__mergeResults()__.

Below, we run `derfinder` for the example data included in the package. The
steps are:

1. Load derfinder
1. Create a directory where we'll store the results
1. Generate the pre-requisites for the models to use with the example data
1. Generate the statistical models
1. Analyze the example data for chr21
1. Merge the results (only one chr in this case, but in practice there'll be more)

```{r loadDerfinder, bootstrap.show.code=TRUE}
## Load derfinder
library('derfinder')

## The output will be saved in the 'report' directory
dir.create('report', showWarnings = FALSE, recursive = TRUE)
```

The following code runs `derfinder`.

```{r runDerfinderFake, eval=FALSE, bootstrap.show.code=TRUE}
## Save the current path
initialPath <- getwd()
setwd(file.path(initialPath, 'report'))

## Generate output from derfinder

## Collapse the coverage information
collapsedFull <- collapseFullCoverage(list(genomeData$coverage), 
verbose=TRUE)

## Calculate library size adjustments
sampleDepths <- sampleDepth(collapsedFull, probs=c(0.5), nonzero=TRUE, 
verbose=TRUE)

## Build the models
group <- genomeInfo$pop
adjustvars <- data.frame(genomeInfo$gender)
models <- makeModels(sampleDepths, testvars=group, adjustvars=adjustvars)

## Analyze chromosome 21
analysis <- analyzeChr(chr='21', coverageInfo=genomeData, models=models, 
cutoffFstat=1, cutoffType='manual', seeds=20140330, groupInfo=group, 
mc.cores=1, writeOutput=TRUE, returnOutput=TRUE)

## Save the stats options for later
optionsStats <- analysis$optionsStats

## Change the directory back to the original one
setwd(initialPath)
```

For convenience, we have included the `derfinder` results as part of 
`regionReport`. Note that the above functions are routinely checked as part
of `derfinder`.


```{r runDerfinderReal, bootstrap.show.code=TRUE}
## Copy previous results
file.copy(system.file(file.path('extdata', 'chr21'), package='derfinder', 
mustWork=TRUE), 'report', recursive=TRUE)
```

Next, proceed to merging the results.

```{r mergeResults, bootstrap.show.code=TRUE, bootstrap.show.message=FALSE}
## Merge the results from the different chromosomes. In this case, there's 
## only one: chr21
mergeResults(chrs = 'chr21', prefix = 'report',
    genomicState = genomicState$fullGenome)
```



### Create report

Once the `derfinder` output has been generated and merged, use
__derfinderReport()__ to create the HTML report.


```{r loadLib, message=FALSE, bootstrap.show.code=TRUE}
## Load derfindeReport
library('regionReport')
```


```{r createReportFake, eval=FALSE, bootstrap.show.code=TRUE}
## Generate the HTML report
report <- derfinderReport(prefix='report', browse=FALSE,
    nBestRegions=15, makeBestClusters=TRUE, outdir='html',
    fullCov=list('21'=genomeDataRaw$coverage), optionsStats=optionsStats)
```


```{r createReportReal, echo=FALSE, message=FALSE, bootstrap.show.code=FALSE}
## Generate the HTML report in a clean environment
library('devtools')
cat("## Generate the report in an isolated environment
## This helps avoids conflicts with generating the vignette
library(derfinder)
library(regionReport)

## Load optionsStats
load(file.path('report', 'chr21', 'optionsStats.Rdata'))

## Create report
report <- derfinderReport(prefix='report', browse=FALSE,
    nBestRegions=15, makeBestClusters=TRUE, outdir='html',
    fullCov=list('21'=genomeDataRaw$coverage), optionsStats=optionsStats)
    
## Clean up
file.remove('derfinderReport-isolated.R')
", file='derfinderReport-isolated.R')
clean_source('derfinderReport-isolated.R', quiet=TRUE)
```

Once the output is generated, you can browse the report from `R` using 
__browseURL()__ as shown below.

```{r vignetteBrowse, eval=FALSE, bootstrap.show.code=TRUE}
## Browse the report
browseURL(report)
```

You can compare the resulting report with the pre-compiled report using the 
following code.

```{r openIncludedReport, eval=FALSE}
browseURL(system.file(file.path('basicExploration', 'basicExploration.html'),  
    package = 'regionReport', mustWork = TRUE))
```

### Notes

Note that the reports require an active Internet connection to render correctly.

The report is self-explanatory and will change some of the text depending on the
input options.

If the report is taking too long to compile (say more than 3 hours), you might
want to consider setting _nBestCluters_ to a small number or even set 
_makeBestClusters_ to `FALSE`.


# Advanced arguments

If you are interested in using the advanced arguments, use `derfinder::advancedArg()` as shown below:

```{r 'advancedArg'}
## URLs to advanced arguemtns
derfinder::advancedArg('derfinderReport', package = 'regionReport',
    browse = FALSE)
## Set browse = TRUE if you want to open them in your browser
```


# Reproducibility

Code for creating the vignette

```{r createVignette, eval=FALSE, bootstrap.show.code=TRUE}
## Create the vignette
library('knitrBootstrap') 

knitrBootstrapFlag <- packageVersion('knitrBootstrap') < '1.0.0'
if(knitrBootstrapFlag) {
    ## CRAN version
    system.time(knit_bootstrap('regionReport.Rmd', chooser=c('boot', 'code'), show_code = TRUE))
    unlink('regionReport.md')
} else {
    ## GitHub version
    library('rmarkdown')
    system.time(render('regionReport.Rmd',
        'knitrBootstrap::bootstrap_document'))
}
## Note: if you prefer the knitr version use:
# library('rmarkdown')
# system.time(render('regionReport.Rmd', 'html_document'))

## Extract the R code
library('knitr')
knit('regionReport.Rmd', tangle = TRUE)

## Copy report output to be distributed with the package for comparison 
## purposes
file.copy(file.path('report', 'html', 'basicExploration.html'),
    '../inst/basicExploration/basicExploration.html', overwrite=TRUE)
       
## Clean up
file.remove('derfinderPlotRef.bib')
#unlink('regionReport_files', recursive=TRUE)
unlink('report', recursive = TRUE)
```

Date the vignette was generated.

```{r vignetteReproducibility1, echo=FALSE}
## Date the report was generated
Sys.time()
```

Wallclock time spent generating the vignette.

```{r vignetteReproducibility2, echo=FALSE, bootstrap.show.code=FALSE}
## Processing time in seconds
totalTimeVignette <- diff(c(startTimeVignette, Sys.time()))
round(totalTimeVignette, digits=3)
```

`R` session information.

```{r vignetteReproducibility3, echo=FALSE}
## Session info
session_info()
```

# Bibliography

This vignette was generated using `knitrBootstrap` `r citep(bib[['hester2014knitrbootstrap']])`
with `knitr` `r citep(bib[['xie2014knitr']])` and `rmarkdown` `r citep(bib[['allaire2014rmarkdown']])` running behind the scenes.

Citations made with `knitcitations` `r citep(bib[[1]])`.

```{r vignetteBiblio, results='asis', echo=FALSE}
## Print bibliography
bibliography()
```
